### 2022-11-14 NSF2022/Description/specificaim2.tex

To briefly describe IE formulations of PDEs, we consider the Stokes
equation~\eqref{eqn:stokes} as an example. Its solution can be written
as a convolution 
\begin{align}
  \uu(\xx) = \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
  \label{eqn:conv}
\end{align}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega(t) \rightarrow \RR^2$
is an unknown density function.
%This is analogous to the problem of
%finding the global electrostatic potential $u$ with surface charge
%density $\sigma$
%\begin{align*}
%  u(x) = -\frac{1}{4\pi}\int_{\partial \Omega(t)} 
%    \frac{1}{|\xx - \yy|}\sigma(y) \, ds_\yy.
%\end{align*}
The IE reformulation has several immediate advantages including a
dimension reduction since $\ssigma$ is defined only on $\partial
\Omega(t)$, and the far field boundary condition of the PDE is
automatically satisfied. The density function $\ssigma$ is chosen to
satisfy the boundary condition, and this results in an IE 
\begin{align}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \partial\Omega(t),
  \label{eqn:bie}
\end{align} where $\gg$ is the boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. For numerical stability,
choices of $G$ that result in second-kind Fredholm integral equations,
meaning $\lambda \neq 0$, are preferred.

As an illustration, the problem of finding the global
electrostatic potential $u$ with surface
charge density $\sigma$ is a special case of \eqref{eqn:conv}, where
$G(\xx - \yy) = (4\pi \epsilon_0)^{-1}|\xx - \yy|^{-1}$
in three dimensions
($-(2\pi \epsilon_0)^{-1}\log|\xx - \yy|$ in two dimensions)
with $\epsilon_0$ the permittivity of vacuum.  In this context, 
~\eqref{eqn:bie} is analogous to Gauss's Law.

@@

\begin{wrapfigure}[]{l}{2.6in}
  \centering
  \includegraphics[width=2.5in]{figures/SA2Figures/confinement.pdf}
  \caption{\label{fig:stenosis} \footnotesize A granule vesicle passing
  through a constricted geometry.}
\end{wrapfigure}

In experiments, it is known that interactions between bounding walls and
granules cannot be ignored~\cite{qua-gan-you2021,
KaouiHartingMisbah2011_PRE, her-sto-gra2005}. Therefore, the necessary
modification to the code will be implemented so that confinement can be
considered. This modification requires introducing an additional layer
potential on the bounding walls, and elements of this modification have
already been incorporated (Figure~\ref{fig:stenosis}). However, to
consider rheological properties of the suspensions, a periodic domain,
such as a Couette device, will be considered. Such multiply-connected
domains require additional modifications to the IE formulation. PI
Quaife has considered suspensions with confinement for several projects
including dynamic suspensions~\cite{qua-bir2014, bys-sha-qua2020} and
porous media~\cite{chi-moo-qua2020, dea-qua-bir-jua2018,
moo-che-chi-qua2022, qua-gan-you2021}.
\begin{wrapfigure}[16]{r}{2.1in}
  \vspace{-8pt}
  \centering
  \input{figures/near_sing.tikz}
  \vspace{-15pt}
  \caption{\label{fig:interpquad} \footnotesize An interpolation-based
  quadrature method that works in two and three dimensions. The
  spectrally-accurate trapezoid rule is used in points in
  $\mathtt{Far}(\Gamma_i)$, while one dimensional interpolant is used
  for points in $\mathtt{Near}(\Gamma_i)$. $h$ is the arclength
  spacing.}
\end{wrapfigure}

@@

Since we are interested in round (2D) and spherical (3D) granules, we
will exploit the fact that integrals can be computed with spectral
accuracy by transforming the integrals to to a Fourier (2D) or spherical
harmonic (3D) basis. This technique can be used to evaluate layer
potentials for both Stokes~\cite{cor-vee2018} and the screened Laplace
equation~\cite{koh-cor-che-vee2021}. However, this quadrature is not
appropriate or other shapes, and an alternative approach is necessary.
The Barycentric-type formulae for Cauchy integrals results in a
uniformly bounded quadrature error, but the method is described only for
the two-dimensional Laplace and Stokes equations. PI Quaife recently
showed that these quadrature methods can be extended to other
two-dimensional PDEs~\cite{che-lin-her-qua2022}. The method uses
singularity subtraction to write the fundamental solution of the PDE as a
sum of two terms: a logarithmic function and a continuously
differentiable function. Then the Barycentric-type formulae can be
applied to the logarithmic function, and the standard trapezoid
rule~\cite{tre-wei2014} can be applied to the other function. These
singularity subtraction methods can also accurately compute derivatives
of layer potentials~\cite{car2021, car2020, car-kha-kim2018} which is
required, for example, when computing the phase
stress~\eqref{eqn:stressbalance}. Singularity subtraction methods can
also be applied in three dimensions, but not combined with
Barycentric-type formulae. Alternatively, smoothed variations of the
fundamental solution can be used to address nearly-singular
integrals~\cite{beale1, cor2001, tlu-bea2019}. Alternatively, efficient
quadrature methods can be applied directly to accurately compute
singular and near-singular integrals~\cite{khayat_2005,
gimbutas2013sisc}.

@@

For a general double-well potential $W(\phi)$ with local minima at
$\phi_0$ and $\phi_1$, we propose to interpret the solution
of~\eqref{eqn:phase} as a root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  F[\phi] = -\epsilon^2 \Delta \phi + W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (F'(\phi^N))^{-1} F(\phi^N)$, where
$F'(\phi^N) = -\epsilon^2 \Delta + W''(\phi^N)$ is the
Fr\'{e}chet derivative of $F$, and $\alpha$ is the step size that will
be determined using backtracking. This requires solving the variable
coefficient PDE $-\epsilon^2 \Delta v + W''(\phi^{N}) v =
F[\phi^N]$, which is challenging with an integral equation since a
fundamental solution is not readily available. However, by choosing a
double-well potential with $W''(\phi_0) = W''(\phi_1) = \beta$, we
propose to instead use an inexact Newton's method where $F^{-1}$ is
approximated with the solution of $-\epsilon^2 \Delta v +
\beta v = F[\phi^N]$. We test the efficacy of the exact and
inexact Newton's method applied to~\eqref{eqn:phaseRoot} in $(0,1)
\subset \RR$ by discretizing $F'$ with a second-order finite difference
stencil. The double-well potential is illustrated in
Figure~\ref{fig:CA}(a), and both the exact and inexact Newton's methods
converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach of a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.

>>>>

To briefly describe IE formulations of PDEs, we consider the Stokes
equation~\eqref{eqn:stokes} as an example. Its solution can be written
as a convolution 
\begin{align}
  \uu(\xx) = \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
  \label{eqn:conv}
\end{align}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega(t) \rightarrow \RR^2$
is an unknown density function. The integral
representation~\eqref{eqn:conv} is known as a layer potential.
%This is analogous to the problem of
%finding the global electrostatic potential $u$ with surface charge
%density $\sigma$
%\begin{align*}
%  u(x) = -\frac{1}{4\pi}\int_{\partial \Omega(t)} 
%    \frac{1}{|\xx - \yy|}\sigma(y) \, ds_\yy.
%\end{align*}
The IE reformulation has several immediate advantages including a
dimension reduction since $\ssigma$ is defined only on $\partial
\Omega(t)$, and the far field boundary condition of the PDE is
automatically satisfied. The density function $\ssigma$ is chosen to
satisfy the boundary condition, and this results in an IE 
\begin{align}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega(t)} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \partial\Omega(t),
  \label{eqn:bie}
\end{align} where $\gg$ is the boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. For numerical stability,
choices of $G$ that result in second-kind Fredholm integral equations,
meaning $\lambda \neq 0$, are preferred.

As an illustration, the problem of finding the global electrostatic
potential $u$ with surface charge density $\sigma$ is a special case of
\eqref{eqn:conv}, where $G(\xx - \yy) = (4\pi \epsilon_0)^{-1}|\xx -
\yy|^{-1}$ in three dimensions ($-(2\pi \epsilon_0)^{-1}\log|\xx - \yy|$
in two dimensions) with $\epsilon_0$ the permittivity of vacuum.  In
this context, equation~\eqref{eqn:bie} is analogous to Gauss's Law.

@@

\begin{wrapfigure}[6]{l}{2.6in}
  \centering
%  \vspace{-10pt}
  \includegraphics[width=2.5in]{figures/SA2Figures/confinement.pdf}
  \vspace{-10pt}
  \caption{\label{fig:stenosis} \footnotesize A granule vesicle passing
  through a constricted geometry.}
\end{wrapfigure}

In experiments, it is known that interactions between bounding walls and
granules cannot be ignored~\cite{qua-gan-you2021,
KaouiHartingMisbah2011_PRE, her-sto-gra2005}. Therefore, the necessary
modification to the code will be implemented so that confinement can be
considered. This modification requires introducing an additional layer
potential on the bounding walls, and elements of this modification have
already been incorporated (Figure~\ref{fig:stenosis}). However, to
consider rheological properties of the suspensions, a periodic domain,
such as a Couette device, will be considered. Such multiply-connected
domains require additional modifications to the IE formulation. PI
Quaife is qualified to perform this task since he has simulated
suspensions with confinement for several projects including dynamic
suspensions~\cite{qua-bir2014, bys-sha-qua2020} and porous
media~\cite{chi-moo-qua2020, dea-qua-bir-jua2018, moo-che-chi-qua2022,
qua-gan-you2021}.
\begin{wrapfigure}[16]{r}{2.1in}
  \vspace{-4pt}
  \centering
  \input{figures/near_sing.tikz}
  \vspace{-15pt}
  \caption{\label{fig:interpquad} \footnotesize An interpolation-based
  quadrature method that works in two and three dimensions. The
  spectrally-accurate trapezoid rule is used in points in
  $\mathtt{Far}(\Gamma_i)$, while one dimensional interpolant is used
  for points in $\mathtt{Near}(\Gamma_i)$. $h$ is the arclength
  spacing.}
\end{wrapfigure}

@@

Since we are interested in round (2D) and spherical (3D) granules, we
will exploit the fact that integrals can be computed with spectral
accuracy by transforming the integrals to to a Fourier (2D) or spherical
harmonic (3D) basis. This technique can be used to evaluate layer
potentials for both Stokes~\cite{cor-vee2018} and the screened Laplace
equation~\cite{koh-cor-che-vee2021}. However, this quadrature is not
appropriate or other shapes, and an alternative approach is necessary.
The Barycentric-type formulae for Cauchy integrals results in a
uniformly bounded quadrature error, but the method is described only for
the two-dimensional Laplace and Stokes equations. PI Quaife recently
showed that these quadrature methods can be extended to other
two-dimensional PDEs~\cite{che-lin-her-qua2022}. The method uses
singularity subtraction to write the fundamental solution of the PDE as a
sum of two terms: a logarithmic function and a continuously
differentiable function. Then the Barycentric-type formulae can be
applied to the logarithmic function, and the standard trapezoid
rule~\cite{tre-wei2014} can be applied to the other function. These
singularity subtraction methods can also accurately compute derivatives
of layer potentials~\cite{car2021, car2020, car-kha-kim2018} which is
required, for example, when computing the phase
stress~\eqref{eqn:stressbalance}. Singularity subtraction methods can
also be applied in three dimensions, but not combined with
Barycentric-type formulae. Alternatively, smoothed variations of the
fundamental solution can be used to address nearly-singular
integrals~\cite{beale1, cor2001, tlu-bea2019}. Finally, efficient
quadrature methods can be applied directly to accurately compute
singular and near-singular integrals~\cite{khayat_2005,
gimbutas2013sisc}.

@@

For a general double-well potential $W(\phi)$ with local minima at
$\phi_0$ and $\phi_1$, we propose to interpret the solution
of~\eqref{eqn:phase} as a root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  F[\phi] = -\epsilon^2 \Delta \phi + W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (F'(\phi^N))^{-1} F(\phi^N)$, where
$F'(\phi^N) = -\epsilon^2 \Delta + W''(\phi^N)$ is the
Fr\'{e}chet derivative of $F$, and $\alpha$ is the step size that will
be determined using backtracking. This requires solving the variable
coefficient PDE $-\epsilon^2 \Delta v + W''(\phi^{N}) v =
F[\phi^N]$, which is challenging with an integral equation since a
fundamental solution is not readily available. However, by choosing a
double-well potential with $W''(\phi_0) = W''(\phi_1) = \beta$, we
propose to instead use an inexact Newton's method where $(F')^{-1}$ is
approximated with the solution of $-\epsilon^2 \Delta v +
\beta v = F[\phi^N]$. We test the efficacy of the exact and
inexact Newton's method applied to~\eqref{eqn:phaseRoot} in $(0,1)
\subset \RR$ by discretizing $F'$ with a second-order finite difference
stencil. The double-well potential is illustrated in
Figure~\ref{fig:CA}(a), and both the exact and inexact Newton's methods
converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach of a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.
